---
title: "Final Project"
output: html_document
date: "2025-02-26"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Load the libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(naniar)
library(ggmap)
library(shiny)
```

```{r}
options(scipen = 999)

locations <- read.csv("Health_Facility_General_Information.csv") %>% 
  clean_names()

cardiac <- read_csv("data/cardiac-surgery.csv") %>% 
  clean_names()
```

```{r}
#COMPARING NUMBER OF TOTAL PROCEDURES WITH AVERAGE MORTALITY RATE
total_cases <- cardiac %>% 
  group_by(facility_id) %>% 
  summarize(total_procedures = sum(number_of_cases), avg_mortality = mean(observed_mortality_rate)) %>%
  ggplot(aes(x = total_procedures, y = avg_mortality))+
  geom_point(color = "olivedrab") +
  labs(title = "Average Mortality Rate compared with Procedure Numbers ",
       x = "Total Number of Procedure",
       y = "Mortality Rate")+
  theme(axis.text.y = element_text(angle = 0, hjust = 1, size = 5))
total_cases
```


```{r}
locations <- locations %>% 
    group_by(facility_id) %>% 
    summarise(across(everything(), first)) 
```

```{r}
cardiac <- left_join(cardiac, locations, by="facility_id")
```
Map:
```{r}
register_stadiamaps("e2de824a-0995-4a51-915c-33c14c061e7b", write = FALSE) 
cardiac %>%  #GET MAX LAT AND LON FROM CODE BELOW!!! EASY MAKES SENSE
  select(facility_longitude, facility_latitude) %>% 
  summary()

lat <- c(40.58, 44.70) #we got tehse values from summary above
long <- c(-78.87, -72.98)
bbox <- make_bbox(long, lat, f = 0.03)

map1 <- get_stadiamap(bbox, maptype = "stamen_terrain", zoom=7) #Make ur basemap, there are other styles which you could find from this link, https://docs.stadiamaps.com/themes/

ggmap(map1)

ggmap(map1) + 
  geom_point(data = cardiac, aes(facility_longitude, facility_latitude, size = number_of_cases)) + #WE just add the chart in without using ggplot! 
  labs(x= "Longitude", y= "Latitude", title="Cardiac Locations") 
```



## Load the Data
```{r}

```

```{r}

```

## Cleaning the Data
```{r}

```

```{r}
cardiac <- merged_data_set %>% 
  mutate(across(everything(), tolower))
```

```{r}
merged_data_set <- merged_data_set %>%
  mutate(across(everything(), tolower)) %>% 
  mutate(across(everything(), ~ str_replace_all((.), " ", "_"))) %>% 
  mutate(across(everything(), ~ str_replace_all((.), "__", "_"))) %>% 
  mutate(across(-"Year of Hospital Discharge", ~ str_remove((.), "-")))
merged_data_set
```

Have to separate the data, as some of them are over a range of a couple of years, some of them are year by year
```{r}
cardiac <- cardiac %>% 
  filter(str_detect(year_of_hospital_discharge, "-", negate = TRUE))
```


## Plots

What is the procedure with the highest observed mortality rate?
```{r}
cardiac %>% 
  group_by(procedure) %>% 
  summarize(avg_mortality = mean(observed_mortality_rate)) %>% 
  arrange(-avg_mortality)
```

```{r}
cardiac %>% 
  group_by(procedure) %>% 
  summarize(avg_mortality = mean(observed_mortality_rate)) %>% 
  ggplot(aes(x = reorder(procedure, avg_mortality), y = avg_mortality))+
  geom_col(color = "black", fill = "lightblue")+
  coord_flip()+
  labs(title = "Average Observed Mortality Rate by Procedure",
       x = "Procedure",
       y = "Mortality Rate")+
  theme_light(base_size = 12)
```


How has the mortality rate changed for cardiac surgery over time in each region?
```{r}
#I'll come back to this later if I figure out the year stuff but feel free to take a crack at it :)
cardiac %>%
  ggplot(aes(x = observed_mortality_rate, y = year_of_hospital_discharge))+
  geom_col(color = "steelblue") + geom_smooth(method = lm, se= TRUE) + 
  labs(title = "Regional Cardiac Surgery Morality Rates",
       x = "Year of Hospital Discharge",
       y = "Observed Mortality Rate")
```


How does the expected mortality rate compare to the observed mortality rate?
```{r}
#I don't know which graph to plot for this one but if you can figure it out that would be great. I was thinking maybe scatter but I don't know for sure

#Expected mortality vs observed mortality

cardiac %>%
  ggplot(aes(x = expected_mortality_rate, y = observed_mortality_rate))+
  geom_point(color = "steelblue") + geom_smooth(method = lm, se= TRUE) + 
  labs(title = "Expected vs Observed Mortality Rate",
       x = "Expected Mortality",
       y = "Observed Mortality")
```


Which region has the highest mortality rate?
```{r}
cardiac %>% 
  group_by(region) %>% 
  summarize(avg_mortality = mean(observed_mortality_rate)) %>% 
  arrange(-avg_mortality)
```

```{r}
cardiac %>% 
  group_by(region) %>% 
  summarize(avg_mortality = mean(observed_mortality_rate)) %>% 
  ggplot(aes(x = reorder(region, avg_mortality), y = avg_mortality))+
  geom_col(color = "black", fill = "lightpink")+
  coord_flip()+
  labs(title = "Average Observed Mortality Rate per Region",
       x = "Region",
       y = "Mortality Rate")+
  theme_light(base_size = 12)
```

It would be cool to plot these hospitals on a map but we would need coordinates for that :,)

## Shiny App
Don't mind the code and names for this, I'm testing out some stuff :D
```{r}
library(shinydashboard)

ui <- dashboardPage(skin = "olivedrab",
                    
  dashboardHeader(title = "Cardiac Dashboard"),
  
  ## Sidebar content
  dashboardSidebar(
    sidebarMenu(
      
      menuItem("Observed Mortality Rate", 
               tabName = "dashboard", 
               icon = icon("dashboard")),
      menuItem("Counts", 
               tabName = "widgets", 
               icon = icon("th")),
      menuItem("Cases",
               tabName = "cases",
               icon = icon("folder-open")),
      menuItem("Hospital Map",
               tabName = "map",
               icon = icon("earth-americas"))
    )
    ),
  
  ## Body content
  dashboardBody(
    
    tabItems(
  
      ## First tab content    
      tabItem(tabName = "dashboard",
        fluidRow(
          box(plotOutput("plot1", height = 250)), # box is a container for the plot
          box(title = "Controls", # box is a container for the controls
              selectInput("region", 
                          "Select Region of Interest:", 
                          choices=unique(cardiac$region))
          )
          )
          ),

      ## Second tab item 
      tabItem(tabName = "widgets",
        fluidRow(
          box(plotOutput("plot2", height = 250)), # box is a container for the plot
          box(title = "Controls", # box is a container for the controls
              radioButtons("x", 
               "Select Fill Variable", 
               choices=c("region", "procedure"),
               selected="region")
          )
          )
          ),
      
      ## Third tab item
      tabItem(tabName = "cases",
        fluidRow(
          box(plotOutput("plot3", height = 250)), # box is a container for the plot
          box(title = "Controls", # box is a container for the controls
              selectInput("y", 
               "Select Hospital", 
               choices=c(unique(cardiac$hospital_name))
          )
          )
          )
          ),
        
        tabItem(tabName = "map",
          fluidRow(
            box(plotOutput("plot4", height = 250)),
            box(title = "Controls",
                selectInput("z",
                            "Select Region",
                             choices = c(unique(cardiac$region))
          
          )
          )
          )
          )
          )
          )
  )
          
server <- function(input, output, session) {
  
  session$onSessionEnded(stopApp)

  output$plot1 <- renderPlot({

    cardiac %>%
      filter(region == input$region) %>%
      ggplot(aes(x = observed_mortality_rate)) + 
      geom_density(color = "black", fill = "steelblue", alpha = 0.6)
  })
  
    output$plot2 <- renderPlot({
    
    cardiac %>% 
      ggplot(aes_string(x="region", fill=input$x))+
      geom_bar(position="dodge", alpha=0.8, color="black")+
      labs(x=NULL, y=NULL, fill="Fill Variable")
    
  })
    
    output$plot3 <- renderPlot({
    
    cardiac %>% 
      filter(hospital_name == input$y) %>% 
      ggplot(aes_string(x="year_of_hospital_discharge", y = "number_of_cases"))+
      coord_flip()+
      geom_col(alpha=0.8, color="black")+
      labs(x=NULL, y=NULL)
    
  })
    
}

shinyApp(ui, server)
```

